---
title: "hw_2"
author: "Stephanie Tsai"
date: "January 25, 2019"
output:
  html_document:
    theme: spacelab
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
---
```{r}
library("tidyverse")
```


## Mammals Sleep
For this assignment, we are going to use built-in data on mammal sleep patterns.  
```{r}
msleep
```

1. From which publication are these data taken from? Don't do an internet search; show the code that you would use to find out in R.
```{r}
?msleep
```



2. Provide some summary information about the data to get you started; feel free to use the functions that you find most helpful.

```{r}
glimpse(msleep)
```
```{r}
names(msleep)
```

##Using filters

3. Make a new data frame focused on body weight, but be sure to indicate the common name and genus of each mammal. Sort the data in descending order by body weight.
```{r}
msleep_filterer <- msleep %>%
  select(name, genus, bodywt) %>%
  arrange(desc(bodywt))


msleep_filterer
```


4. We are interested in two groups; small and large mammals. Let's define small as less than or equal to 1kg body weight and large as greater than or equal to 200kg body weight. For our study, we are interested in body weight and sleep total Make two new dataframes (large and small) based on these parameters. Sort the data in descending order by body weight.

##Small Mammals
```{r}
small_animal <- msleep %>%
  select(name, genus, bodywt, sleep_total) %>%
  filter(bodywt <= 1) %>%
  arrange(desc(bodywt))

small_animal
```

## Large Mammals
```{r}
large_animal <- msleep %>%
  select(name, genus, bodywt, sleep_total) %>%
  filter(bodywt >= 200) %>%
  arrange(desc(bodywt))
  
large_animal
```

##Averages

5. Let's try to figure out if large mammals sleep, on average, longer than small mammals. What is the average sleep duration for large mammals as we have defined them?
```{r}
#mean(large$sleep_total)

mean_largesleep <- mean(large_animal$sleep_total)
mean_largesleep
```

6. What is the average sleep duration for small mammals as we have defined them?
```{r}
mean_smallsleep <-mean(small_animal$sleep_total)
mean_smallsleep
```
```{r}
#Finding sleep duration
if(mean_smallsleep > mean_largesleep){
  print("small animals sleep longer")
} else if(mean_largesleep > mean_smallsleep){
  print("large animals sleep longer")
} else {
  print("No difference in sleep duration")
}
  
```

##Animals sleeping at least 18 hrs per day

7. Which animals sleep at least 18 hours per day? Be sure to show the name, genus, order, and sleep total. Sort by order and sleep total.
```{r}
#Organized by decending sleep total
long_sleep <- msleep %>%
    select(name, genus, order, sleep_total) %>%
    filter(sleep_total >= 18) %>%
    arrange(desc(sleep_total))

long_sleep
    
```
```{r}
##organized by order name
order_sleep <- msleep %>%
    select(name, genus, order, sleep_total) %>%
    filter(sleep_total >= 18) %>%
    arrange(order)

order_sleep
```

```{r}
##Organized by order and sleep total
complete_sleep <- msleep %>%
  select(name, order, genus, sleep_total) %>%
  filter(sleep_total >= 18) %>%
  arrange(order, sleep_total)

complete_sleep
```


