---
title: "lab4_hw"
author: "Stephanie Tsai"
date: "February 8, 2019"
output: 
  html_document: 
    keep_md: yes
---

```{r}
library(tidyverse)
library(skimr)
```


```{r}
life_history <- readr::read_csv("C:/Users/Stephanie T/Desktop/Master File/class_files-master/mammal_lifehistories_v2.csv")
```

Rename some of the variables. Notice that I am replacing the old `life_history` data.
```{r}
life_histories <- 
  life_history %>% 
  rename( genus = Genus,
          wean_mass    = `wean mass`,
          max_life     = `max. life`,
          litter_size  = `litter size`,
          litters_yr   = `litters/year`
          )

life_histories
```

1. Explore the data using the method that you prefer. Below, I am using a new package called `skimr`. If you want to use this, make sure that it is installed.


```{r}
life_histories %>% 
  skimr::skim()
```

2. Run the code below. Are there any NA's in the data? Does this seem likely?
```{r}
life_histories %>% 
  summarize(number_nas= sum(is.na(life_history)))
```
```{r}
#The summarize function states there are no NA data in the data frame, but this is highly unlikely because our previous code shows that there are -999 (usually indicates an NA from excel)
```



3. Are there any missing data (NA's) represented by different values? How much and where? In which variables do we have the most missing data? Can you think of a reason why so much data are missing in this variable?
```{r}
#Values used to represent NA: -999

# These values are in all the numeric data and integer data
```


```{r}
life_histories_na <- 
  life_histories %>%
  na_if("-999")

life_histories_na %>% 
  purrr::map_df(~ sum(is.na(.))) %>%
  gather(variable, value = "num_NA") %>%
  arrange(desc(num_NA))
```
```{r}
# The variable with the most missing data is in wean_mass at 1039 NAs. It probably is difficult to determine exactly when an animal has entered the weaning phase, so not all animals could be weighed at the time.
```


4. Compared to the `msleep` data, we have better representation among taxa. Produce a summary that shows the number of observations by taxonomic order.
```{r}
taxa_summary <- life_histories_na %>% 
  group_by(order) %>% 
  summarize(n_total = n())

taxa_summary
  
```


5. Mammals have a range of life histories, including lifespan. Produce a summary of lifespan in years by order. Be sure to include the minimum, maximum, mean, standard deviation, and total n.
```{r}
#Adjustments to make: max lifespan to years from months
# Reference http://esapubs.org/archive/ecol/E084/093/
```
## Tidy data
```{r}

summary_life <- life_histories_na %>%
  mutate(yr_max_life = max_life/12) %>%
  group_by(order) %>%
  summarise(min_lifespan_yrs = min(yr_max_life, na.rm = TRUE),
   max_lifespan_yrs = max(yr_max_life, na.rm = TRUE),
   mean_lifespan_yrs = mean(yr_max_life, na.rm = TRUE),
   sd_lifespan_yrs = sd(yr_max_life, na.rm = TRUE),
   n_total = n()) 

summary_life
```




6. Let's look closely at gestation and newborns. Summarize the mean gestation, newborn mass, and weaning mass by order. Add a new column that shows mean gestation in years and sort in descending order.  

## Make initial summary chart
```{r}
#Note mass is by grams, gestation is by months
life_newBorns <- life_histories_na %>%
  select(order, gestation, newborn, wean_mass) %>%
  group_by(order) %>%
  summarise(mean_gestation_months = mean(gestation),
            mean_gestation_yrs = mean(gestation/12),
            mean_newborn_mass = mean(newborn),
            mean_weanmass = mean(wean_mass)) %>%
  arrange(desc(mean_gestation_yrs))

life_newBorns

```

## Which group has the longest mean gestation? What is the common name for these mammals?
  Proboscidea has the longest mean gestation of about 1.77 years. Their common name is elephants.

## Wrap-up
Please review the learning goals and be sure to use the code here as a reference when completing the homework.

A) Work on making a repository for team data
B) Upload data