---
title: "lab3_hw"
author: "Stephanie Tsai"
date: "February 1, 2019"
output: 
  html_document: 
    keep_md: yes
---
```{r}
library(tidyverse)
```

```{r}
getwd()
```

##Loading the data
```{r}
fisheries <- readr::read_csv(file = "C:/Users/Stephanie T/Desktop/class_files-master/class_files-master/FAO_1950to2012_111914.csv") 
```

1. Do you see any potential problems with the column names? Does the error message now make more sense?  
*Column names are duplicated so R manipulated them to unique names*
```{r}
#Column names that are the same risk being considered the same even when the data is supposed to be different. R is trying to fix this by relabeling them as an unique, separate column.
```
# Viewing available data
```{r}
glimpse(fisheries)
```
## Viewing data and Renaming
2. The `make.names()` command is helpful when there are issues with column names. Notice that although the names are still cumbersome, much of the problemtatic syntax is removed.
```{r}
names(fisheries) = make.names(names(fisheries), unique=T) #changes the column names
names(fisheries)

#using make.names() takes the fisheries columns and check if the name is unique and an acceptable name. If not, R will change it to an accepted notation.

#In this case, R changed numeric dates into characters leading with X and removed paranthesis
```


3. Let's rename the columns. Use `rename()` to adjust the names as follows. Double check to make sure the rename worked correctly. Make sure to replace the old fisheries object with a new one so you can keep the column names.
+ country     = Country..Country.  
+ commname    = Species..ASFIS.species.  
+ sciname     = Species..ASFIS.species._2  
+ spcode      = Species..ASFIS.species._1  
+ spgroup     = Species..ISSCAAP.group.  
+ spgroupname = Species..ISSCAAP.group._1  
+ region      = Fishing.area..FAO.major.fishing.area.  
+ unit        = Measure..Measure.  

```{r}
fisheries_2 <- fisheries %>%
  dplyr::rename(
     country     = Country..Country. , 
 commname    = Species..ASFIS.species.  ,
 sciname     = Species..ASFIS.species._2  ,
 spcode      = Species..ASFIS.species._1  ,
 spgroup     = Species..ISSCAAP.group.  ,
 spgroupname = Species..ISSCAAP.group._1 , 
 region      = Fishing.area..FAO.major.fishing.area.  ,
 unit        = Measure..Measure.
  )

fisheries_2
```

## Evaluating Tidyness
4. Are these data tidy? Why or why not, and, how do you know?

```{r}
# The data is organized in a way that makes sense to a person looking at it, but not for R. All the data points will still be read as vectors rather than being specifically grouped to match with its appropriate label. Each year variable does not have a defined observation linked to it, so the numbers associated to year will be read columnarly by R.

```




## Tidy data with gather()

5. We need to tidy the data using `gather()`. The code below will not run because it is commented (#) out. I have added a bit of code that will prevent you from needing to type in each year from 1950-2012 but you need to complete the remainder `QQQ` and remove the `#`.
```{r}
fisheries_tidy <- 
  fisheries_2 %>% 
  gather(num_range('X',1950:2012), key= "year", value= "catch")

fisheries_tidy
```

6. Use `glimpse()` to look at the categories of the variables. Pay particular attention to `year` and `catch`. What do you notice?  

```{r}
glimpse(fisheries_tidy)
```

```{r}
# The year and catch are represented by characters. We would expect them to be numbers. The catch variable has a lot of missing data.

```



7. From question 6 you should see that there are a lot of entries that are missing. In R, these are referred to as NA's but they can be coded in different ways by the scientists in a given study. In order to make the data tidy, we need to deal with them. As a preview to our next lab, run the following code by removing the `#`. It removes the 'X' from the years and changes the `catch` column from a character into a numeric. This forces the blank entries to become NAs. The error "NAs introduced by coercion" indicates their replacement.
```{r}
fisheries_tidy <- 
  fisheries_tidy %>% 
  mutate(
    year= as.numeric(str_replace(year, 'X', '')), #remove the X from year
    catch= as.numeric(str_replace(catch, c(' F','...','-'), replacement = '')) #replace character strings with NA
    )

fisheries_tidy
```

8. Are the data tidy? Why?  

```{r}
# The data is more tidy than before because now R knows that the year column contains numeric years and it is paired to the catch column as one variable to one data point.
```


9. You are a fisheries scientist studying cephalopod catch during 2008-2012. Identify the top five consumers (by country) of cephalopods (don't worry about species for now). Restrict the data frame only to our variables of interest.

##Restrict year to 2008-2012
```{r}
year_narrow <- fisheries_tidy %>%
  select(country, year, spgroupname,commname, catch) %>%
  filter(year > 2008 & year < 2012 | year == 2008 |year == 2012)

year_narrow
```

##Who consumes cephalopods?
```{r}
cephalopod_consumption <- year_narrow %>%
  filter(spgroupname == "Squids, cuttlefishes, octopuses") %>%
  filter(catch >= 1)
  
cephalopod_consumption
```
##Gathering countries
```{r}
#Calculates based on the mean and max catch of cephalopods per country out of all their years

top_consumers <- cephalopod_consumption %>%
  group_by(country) %>%
  summarize(max_Catch = max(catch),
            mean_Catch = mean(catch)) %>%
  arrange(desc(mean_Catch)) #arranged by the average cephalopod catch

top_consumers
```
##The top 5 consumers of cephalopods are Vietnam, China, Peru, Chile, and Japan.


10. Let's be more specific. Who consumes the most `Common cuttlefish`? Store this as a new object `cuttle`.

```{r}
cuttle <- year_narrow %>%
  filter(commname == "Common cuttlefish") %>%
  group_by(country) %>%
  summarize(max_Catch = max(catch),
            mean_Catch = mean(catch)) %>%
  arrange(desc(mean_Catch)) #arranged by the average cephalopod catch
  
cuttle
  
```

##Tunisia consumes the most Common Cuttlefish.

