---
title: "hw_5"
author: "Stephanie Tsai"
date: "February 15, 2019"
output: 
  html_document: 
    keep_md: yes
---
```{r}
library(tidyverse)
library(skimr)
```
## Mammals Life History
Let's revisit the mammal life history data to practice our `ggplot` skills. Some of the tidy steps will be a repeat from the homework, but it is good practice. The [data](http://esapubs.org/archive/ecol/E084/093/) are from: *S. K. Morgan Ernest. 2003. Life history characteristics of placental non-volant mammals. Ecology 84:3402.*

1. Load the data.
```{r}
life_history <- read_csv("C:/Users/Stephanie T/Desktop/FRS_417/class_files-master/mammal_lifehistories_v2.csv")
```


2. Use your preferred function to have a look. Do you notice any problems?
```{r}
skimr::skim(life_history)
```
```{r}
# There are some NAs in the data because there are -999 values
```


3. There are NA's. How are you going to deal with them?
```{r}
life_history_identNA <- 
  life_history %>% 
  na_if("-999") #replace -999 data with NA

life_history_identNA
```
```{r}
skim(life_history_identNA)
```


4. Where are the NA's? This is important to keep in mind as you build plots.
```{r}
# The NAs are all in the numeric data. It is not in the order, family, genus, and species.
```


5. Some of the variable names will be problematic. Let's rename them here as a final tidy step.

```{r}
tidy_lifeHistory <- life_history_identNA %>%
rename(genus = 'Genus',
          wean_mass = 'wean mass',
          max_life = 'max. life',
          litter_size = 'litter size',
          litters_yr = 'litters/year')

names(tidy_lifeHistory)
```


##`ggplot()`
For the questions below, try to use the aesthetics you have learned to make visually appealing and informative plots. Make sure to include labels for the axes and titles.

options(scipen=999)  #cancels the use of scientific notation for the session

6. What is the relationship between newborn body mass and gestation? Make a scatterplot that shows this relationship. 
```{r}
options(scipen=999)

tidy_lifeHistory %>% 
  ggplot(aes(x = newborn, y = gestation)) +
  geom_point()
```

7. You should notice that because of the outliers in newborn mass, we need to make some changes. We didn't talk about this in lab, but you can use `scale_x_log10()` as a layer to correct for this issue. This will log transform the y-axis values.
```{r}
tidy_lifeHistory %>% 
  ggplot(aes(x = newborn, y = gestation)) +
  scale_x_log10() +
  geom_point()
```



