---
title: "hw_6"
author: "Stephanie Tsai"
date: "March 1, 2019"
output: 
  html_document: 
    keep_md: yes
---

## Load the libraries
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(skimr)
library("RColorBrewer")
library("gapminder")
```

## Resources
The idea for this assignment came from [Rebecca Barter's](http://www.rebeccabarter.com/blog/2017-11-17-ggplot2_tutorial/) ggplot tutorial so if you get lost go have a look. Please do not copy and paste her code!  

## Gapminder
For this assignment, we are going to use the dataset [gapminder](https://cran.r-project.org/web/packages/gapminder/index.html). Gapminder includes information about economics, population, and life expectancy from countries all over the world. You will need to install it before use.
```{r}
#install.packages("gapminder")
```


Please load the data into a new object called gapminder.
```{r}
gapminder <- 
  gapminder::gapminder
```

1. Explore the data using the various function you have learned. Is it tidy, are there any NA's, what are its dimensions, what are the column names, etc.
```{r}
#Finding NAs - no NA found
skim(gapminder)
```

```{r}
#Column names
names(gapminder)
```
```{r}
#Dimensions are 1704 rows and 6 columns
dim(gapminder)
```





2. We are interested in the relationship between per capita GDP and life expectancy; i.e. does having more money help you live longer on average. Make a quick plot below to visualize this relationship.
```{r}
gapminder %>% 
  ggplot(aes(x = gdpPercap, y = lifeExp)) +
  geom_point(col = "blue", size = 0.8) +
  labs(title = "Relationship between Life Expectancy and Per Capita GDP",
       y = "Life Expectancy",
       x = "Per Capita GDP") +
  theme_bw()
  
```


3. There is extreme disparity in per capita GDP. Rescale the x axis to make this easier to interpret. How would you characterize the relationship?
```{r}
gapminder %>% 
  ggplot(aes(x = gdpPercap, y = lifeExp)) +
  xlim(0, 7000) +
  geom_point(col = "blue", size = 0.8) +
  labs(title = "Relationship between Life Expectancy and Per Capita GDP",
       y = "Life Expectancy",
       x = "Per Capita GDP") +
  theme_bw()
```

```{r}
# It looks like the higher the GDP then the greater the life expectancy.
```



4. This should look pretty dense to you with significant overplotting. Try using a faceting approach to break this relationship down by year.
```{r}
gapminder %>% 
  ggplot(aes(x = gdpPercap, y = lifeExp)) +
  xlim(0, 7000) +
  geom_point(col = "blue", size = 0.8) +
  labs(title = "Relationship between Life Expectancy and Per Capita GDP",
       y = "Life Expectancy",
       x = "Per Capita GDP") +
  facet_wrap(~year) +
  theme_bw()
```


5. Simplify the comparison by comparing only 1952 and 2007. Can you come to any conclusions?
```{r}
gapminder %>% 
  filter(year == 1952 | year == 2007) %>%  
  ggplot(aes(x = gdpPercap, y = lifeExp)) +
  xlim(0, 7000) +
  geom_point(col = "blue") +
  labs(title = "Relationship between Life Expectancy and Per Capita GDP",
       y = "Life Expectancy",
       x = "Per Capita GDP") +
  facet_wrap(~year) +
  theme_bw()
```
```{r}
#Life expectancy appears to be higher with a higher GDP. The range of per capita GDP did not change much in 2007 compared to 1952, but it looks like it shifted up. 
```


6. Let's stick with the 1952 and 2007 comparison but make some aesthetic adjustments. First try to color by continent and adjust the size of the points by population. Add `+ scale_size(range = c(0.1, 10), guide = "none")` as a layer to clean things up a bit.
```{r}
gapminder %>% 
  filter(year == 1952 | year == 2007) %>%  
  ggplot(aes(x = gdpPercap, y = lifeExp, color = continent, size = pop)) +
  xlim(0, 7000) +
  geom_point(alpha = 0.6) +
  scale_size(range = c(0.1, 10), guide = "none") +
  labs(title = "Relationship between Life Expectancy and Per Capita GDP",
       y = "Life Expectancy",
       x = "Per Capita GDP") +
  facet_wrap(~year) +
  theme_bw()
```


7. Although we did not introduce them in lab, ggplot has a number of built-in themes that make things easier. I like the light theme for these data, but you can see lots of options. Apply one of these to your plot above.
```{r}
gapminder %>% 
  filter(year == 1952 | year == 2007) %>%  
  ggplot(aes(x = gdpPercap, y = lifeExp, color = continent)) +
  xlim(0, 7000) +
  geom_point(alpha = 0.6, aes(size = pop)) +
  scale_size(range = c(0.1, 10), guide = "none") +
  labs(title = "Relationship between Life Expectancy and Per Capita GDP",
       y = "Life Expectancy",
       x = "Per Capita GDP") +
  facet_wrap(~year) +
  theme_light()

#how to add legend for pop?
```

8. What is the population for all countries on the Asian continent in 2007? Show this as a barplot.
```{r}
gapminder %>% 
  filter(continent == "Asia" & year == 2007) %>% 
  ggplot(aes(x = country, y = pop)) +
  geom_bar(stat = "identity", fill = "orchid2") +
  labs(title = "Population by country in Asia (2007)",
       y = "Population",
       x = "Country") +
  coord_flip() +
  theme_bw()
```


9. You should see that China's population is the largest with India a close second. Let's focus on China only. Make a plot that shows how population has changed over the years.
```{r}
gapminder %>% 
  filter(country == "China") %>% 
  ggplot(aes(x = year, y = pop)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  labs(title = "Population Growth in China over time",
       y = "Population",
       x = "Year") +
  theme_bw()
```


10. Let's compare China and India. Make a barplot that shows population growth by year using `position=dodge`. Apply a custom color theme using RColorBrewer.
```{r}
brewer.pal.info
```



```{r}
gapminder %>% 
  filter(country == "China"| country == "India") %>% 
  ggplot(aes(x = year, y = pop, fill = country)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Population Growth over time in China and India",
       y = "Population",
       x = "Year") +
  scale_fill_brewer(palette = "Set1") +
  theme_bw()
```


## Push your final code to [GitHub](https://github.com/FRS417-DataScienceBiologists)
Make sure that you push your code into the appropriate folder. Also, be sure that you have check the `keep md` file in the knit preferences.